Empirical study of land surface effects on local rainfall

### 2017-11-08 chap3_paper_new.tex
add files

\begin{document}
\SweaveOpts{concordance=TRUE}

@@

<<mapping,eval=F, echo = F, results = hide>>=

library(plotKML)
library(raster)
library(rgeos)
library(maptools)
library(fields)
library(RColorBrewer)

## define the boundary of sub-regions
  e1 <- extent(c(1324304.091,1723901.773,-222556.668,80451.265))
  # 400 km by 303 km Qld
  e2 <- extent(c(1346395.141,1534425.527,-1169402.246,-607644.171))
  # 188 km by 561 km NSW
  e11 <- extent(c(145.425, 149.025, -28.475, -25.175))
  e22 <- extent(c(146.675, 149.325, -37.075, -34.675))

  e111 <- extent(c(145.825, 148.625, -28.075, -25.575)) #as(tc_r1,"SpatialPolygons")
  # 306 km by 273.5 km
  e222 <- extent(c(147.075, 148.925, -36.675, -35.075)) #as(tc_r2,"SpatialPolygons")
  # 203 km by 175 km
## the selected regions for mapping

  aus_reg <- readShapePoly("Data\\Geo\\aus_state2011\\simplify_state11_uproj.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
  aus_reg2 <- gSimplify(aus_reg,0.01,topologyPreserve=T)

  mdb_reg <- readShapePoly("Data\\Geo\\MDB\\mdb.shp",proj4string=CRS("+proj=lcc +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0"))
mdb_reg2 <- gSimplify(mdb_reg,0.01,topologyPreserve=TRUE)

  remove(list=c("aus_reg","mdb_reg"))

## Burned area map

fire_0301 <- raster("Data\\Other\\bushfires\\Burnt_Area\\tmpA2003001.burndate.tif")
fire_0301nsw <- crop(fire_0301,e2)
fire_loc <- crop(reproject(fire_0301nsw,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84")),e222)
remove(list=c("fire_0301","fire_0301nsw"))

  pdf("bushfire_nswvic.pdf")
#  jpg("bushfire_nswvic.jpg")
  par(mar=c(3,3,1,6),xpd=T)
  plot(e222,border="white")  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=1.2)
  text(148.4, -36,"Kosciuszko",font=2,cex=0.9)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=1.2)
  text(148, -35.5,"NSW",font=2,cex=1.2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.03,legend.width=1.3,legend.args=list(text="",line=0),lwd.poly=0.1)
legend(149,-36.1,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
 dev.off()

@

<<treeCover,eval=F , echo = F, results = hide>>=

## trend of evi (MODIS product) from GA and ABARES

## tree cover data
  mod44_tc <- brick("Data\\Satellite\\MOD44\\B\\mod44_tc_aus.tif")
  tm <- seq(as.Date("2000-03-06"), as.Date("2010-03-06"), "year")
  mod44_tc <- setZ(mod44_tc, tm, "year")

## tree cover in sub-region
  mod_tc_r11 <- crop(mod44_tc,e1)
  mod_tc_r111 <- reproject(mod_tc_r11,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
  mod_tc_r1 <- crop(mod_tc_r111,e111)

  mod_tc_r22 <- crop(mod44_tc,e2)
  mod_tc_r222 <- reproject(mod_tc_r22,CRS=CRS("+proj=longlat +lat_1=-18 +lat_2=-36 +lat_0=-27 +lon_0=132 +x_0=0 +y_0=0 +ellps=WGS84"))
 mod_tc_r2 <- crop(mod_tc_r222,e222)

 remove(list=c("mod44_tc","mod_tc_r11","mod_tc_r111","mod_tc_r22","mod_tc_r222"))

## apply t-test to tree cover data
  fun <- function(x,m) {
    var_x <- sd(x)
    mu_1 <- mean(x[1:3])
    if (m==1) {y <- x[6:11]} else {y <- x[4:11]}   
    mu_2 <- mean(y)
    z <- (mu_2-mu_1)/var_x
    if (!is.na(z) & abs(z)>qt(0.1,length(z))) {return(y[1]-x[3])} else {return(NA)}
  }

  mod_tc_r1.z <- calc(mod_tc_r1,function(x) fun(x,1))
  mod_tc_r2.z <- calc(mod_tc_r2,function(x) fun(x,2))

  remove(list=c("mod_tc_r1","mod_tc_r2"))

## create plots
  plotclr <- brewer.pal(6,"PiYG")
#"#C51B7D" "#E9A3C9" "#FDE0EF" "#E6F5D0" "#A1D76A" "#4D9221"

  pdf("tc.pdf", width=10, height= 5)
  par(mar=c(3,3,1,7), mfrow=c(1,2))
  plot(e111)
  image(mod_tc_r1.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e111,add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  image.plot(zlim=c(-3,3),breaks=c(-3,-2,-1,-0.2,0.2,1,2,3),lab.breaks=c(-70,-50,-20,"no tree","0 or",20,50,70),legend.only=T,horizontal=F,col=c("#C51B7D","#E9A3C9","#FDE0EF","white","#E6F5D0","#A1D76A","#4D9221"),legend.shrink=0.6,legend.width=0.7,legend.args=list(text=expression(paste(Delta,"tree (% ground)",sep="")),line=0.5,cex=0.9,font=2),axis.args=list(cex.axis=0.8),graphics.reset=F)
  text(146.0, -25.6, "(a)", font=2)
  plot(e222)
  image(mod_tc_r2.z,col=plotclr,breaks=c(-70,-50,-20,0,20,50,70),legend=F,add=T)
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  text(147.2, -35.1, "(b)", font=2)
  dev.off()

pdf("map.pdf")
par(mar=c(1.5,1.5,1.5,1.5))
plot(mdb_reg2,border="grey",lwd=3)
plot(aus_reg2,border="grey",add=T)
legend(150.2,-36,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
#invisible(text(getSpPPolygonsLabptSlots(aus_reg2), labels=as.character(aus_reg2$NAME), cex=0.7,adj=0.5))
box(lty=1)
plot(e111,col="red",add=T)
plot(e222,col="red",add=T)
text(147.2,-26.8,"QLD",cex=1.5)
text(148,-34.7,"NSW/VIC",cex=1.5)
dev.off()

@

@@

<<climaticIndex,eval=F , echo = F, results = hide>>=

library(reshape)
library(chron)
library(fields)
library(rgdal)
library(raster)
library(maptools)
library(graphics)
library(zoo)
library(RColorBrewer) # creates nice color schemes
# library(classInt)     # finds class intervals for continuous variables

library(mgcv)
library(ncdf4)
library(nnet)

## climatic indicator data

file1 <- "Data\\BoM\\SOI\\SOI.csv"
SOI <- read.csv(file1,sep=",",skip=6)
SOI <- melt(SOI,id="Year")
names(SOI) <- c("Year","Month","SOI")
SOI <- SOI[order(SOI$Year,SOI$Month),]
SOI_ts <- ts(SOI$SOI,start=c(1876,1),end=c(2014,4),frequency=12)
SOI_100a <- window(SOI_ts,start=c(1900,1),end=c(2008,12))
SOI_30a <- window(SOI_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SOI
SOI_100 <- stl(SOI_100a,"period")$time.series[,3]+stl(SOI_100a,"period")$time.series[,2]
SOI_30 <- stl(SOI_30a,"period")$time.series[,3]+stl(SOI_30a,"period")$time.series[,2]

file2 <- "Data\\BoM\\SST\\SST34.cdf"
SST34 <- nc_open(file2)
SST34 <- ncvar_get(SST34,SST34$var[[1]])
SST34_ts <- ts(SST34,start=c(1856,1),end=c(2014,4),frequency=12)
SST34_30a <- window(SST34_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST34
SST34_30 <- stl(SST34_30a,"period")$time.series[,3]+stl(SST34_30a,"period")$time.series[,2]

file3 <- "Data\\BoM\\SST\\SST3.cdf"
SST3 <- nc_open(file3)
SST3 <- ncvar_get(SST3,SST3$var[[1]])
SST3_ts <- ts(SST3,start=c(1856,1),end=c(2014,4),frequency=12)
SST3_30a <- window(SST3_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST34
SST3_30 <- stl(SST3_30a,"period")$time.series[,3]+stl(SST3_30a,"period")$time.series[,2]

file4 <- "Data\\BoM\\SST\\SST4.cdf"
SST4 <- nc_open(file4)
SST4 <- ncvar_get(SST4,SST4$var[[1]])
SST4_ts <- ts(SST4,start=c(1856,1),end=c(2014,5),frequency=12)
SST4_30a <- window(SST4_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised SST4
SST4_30 <- stl(SST4_30a,"period")$time.series[,3]+stl(SST4_30a,"period")$time.series[,2]

file5 <- "Data\\BoM\\PDO\\PDO.latest_1900-.txt"
pdo_mth <- read.csv(file5,sep="",header=T)
pdo_mth <- melt.data.frame(pdo_mth,id="YEAR")
names(pdo_mth) <- c("Year","Month","PDO")
pdo_mth <- pdo_mth[order(pdo_mth$Year,pdo_mth$Month),]
PDO_ts <- ts(pdo_mth$PDO,start=c(1900,1),end=c(2014,4),frequency=12)
PDO_100a <- window(PDO_ts,start=c(1900,1),end=c(2008,12))
## deseasonalised PDO
PDO_100 <- stl(PDO_100a,"period")$time.series[,3]+stl(PDO_100a,"period")$time.series[,2]

file7 <- "Data\\BoM\\IOD\\iod_jamstec.txt"
IOD <- read.csv(file7,sep="",header=T)
IOD <- IOD[,4]
IOD_ts <- ts(IOD,start=c(1870,1),frequency=12)
IOD_30a <- window(IOD_ts,start=c(1979,1),end=c(2008,12))
## deseasonalised IOD
IOD_30 <- stl(IOD_30a,"period")$time.series[,3]+stl(IOD_30a,"period")$time.series[,2]
# end of read data

@

@@

<<rainData,eval=F , echo = F, results = hide>>=

## 100 years awap data & 30 years
qld_100 <- read.csv("Data\\BOM\\rainfall\\QLD_month_rf.csv",header=T)
qld_30 <- subset(qld_100,as.Date(qld_100[,1],"%d/%m/%Y")>="1979-01-01" & as.Date(qld_100[,1],"%d/%m/%Y")<="2008-12-01")
qld100_mean <- ts(rowMeans(qld_100[,2:2801],na.rm=T,dims=1),start=c(1900,1),frequency=12)
qld30_mean <- window(qld100_mean,start=c(1979,1),end=c(2008,12))

nsw_100 <- read.csv("Data\\BOM\\rainfall\\NSW_month_rf.csv",header=T)
nsw_30 <- subset(nsw_100,as.Date(nsw_100[,1],"%d/%m/%Y")>="1979-01-01" & as.Date(nsw_100[,1],"%d/%m/%Y")<="2008-12-01")
nsw100_mean <- ts(rowMeans(nsw_100[,2:1185],na.rm=T,dims=1),start=c(1900,1),frequency=12)
nsw30_mean <- window(nsw100_mean, start=c(1979,1),end=c(2008,12))

require(MASS)
fitdistr(qld30_mean,"gamma")  ## assume shape=1
fitdistr(nsw30_mean,"gamma")  ## assume shape =8
ks.test(qld30_mean,1,0.03)
ks.test(nsw30_mean,2.4,0.03)

pdf("hist_rain.pdf", width=8, height = 5)
par(mfrow=c(1,2))
 hist(qld30_mean,main="",xlab=expression(paste("rainfall (mm mont",h^-1,")",sep="")))
 text(200,120,"(a)", font=2)
 hist(nsw30_mean,main="",xlab=expression(paste("rainfall (mm mont",h^-1,")",sep="")))
text(200,60,"(b)", font=2)
dev.off()

## autocorrelation in rainfall
pdf("autocorrelation.pdf")
par(mfrow=c(2,2))
  acf(qld30_mean,lag.max=12,main="QLD")
  acf(nsw30_mean,lag.max=12,main="NSW & VIC")

  pacf(qld30_mean,lag.max=12,main="")
  pacf(nsw30_mean,lag.max=12,main="")  
dev.off()

## deseasonalise & detrend rainfall: assume the deseasonalised monthly zonal rainfall is normal distributed and can be used in pearson cross-correlation function
## use stl function in R
  qld30_deseason <- stl(qld30_mean,s.window=length(qld30_mean),t.window=length(qld30_mean))$time.series[,3]
  nsw30_deseason <- stl(nsw30_mean,s.window=length(nsw30_mean),t.window=length(nsw30_mean))$time.series[,3]
  qld100_deseason <- stl(qld100_mean,s.window=length(qld100_mean),t.window=length(qld100_mean))$time.series[,3]
  nsw100_deseason <- stl(nsw100_mean,s.window=length(nsw100_mean),t.window=length(nsw100_mean))$time.series[,3]

pdf("mean_ts.pdf",width=8,height=4)
par(cex.lab=1,cex.axis=1,mfrow=c(1,2))
plot(qld30_deseason,xlab="Year",ylab="monthly rainfall")
abline(v=2003,col="red",lwd=2)
text(2004.6,150,"2003",cex=1.2,col="red")
text(1980.6,200,"(a)",font=2, cex=1.2)
plot(nsw30_deseason,xlab="Year",ylab="monthly rainfall")
abline(v=2003,col="red",lwd=2)
text(2004.6,120,"2003",cex=1.2,col="red")
text(1980.6,150,"(b)", font=2,cex=1.2)
dev.off()

# Figure 5
pdf("cor_qld.pdf", width = 8, height = 8)
par(mfrow=c(2,3),mar=c(5,4,1,1))
ccf(as.vector(SOI_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST34_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO3.4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST4_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100),as.vector(qld100_deseason),lag.max=24,xlim=c(-24,0),xlab="PDO Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100*SOI_100),as.vector(qld100_deseason),lag.max=12,xlim=c(-12,0),xlab="PDO*SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(IOD_30),as.vector(qld30_deseason),lag.max=12,xlim=c(-12,0),xlab="IOD  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
dev.off()

pdf("cor_nswvic.pdf",width = 8, height = 8)
par(mfrow=c(2,3),mar=c(5,4,1,1))
ccf(as.vector(SOI_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST34_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO3.4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(SST4_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="NINO4  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100),as.vector(nsw100_deseason),lag.max=24,xlim=c(-24,0),xlab="PDO Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(PDO_100*SOI_100),as.vector(nsw100_deseason),lag.max=12,xlim=c(-12,0),xlab="PDO*SOI  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
ccf(as.vector(IOD_30),as.vector(nsw30_deseason),lag.max=12,xlim=c(-12,0),xlab="IOD  Lag",ylab="CCF",cex.axis=1.2,cex.lab=1.3)
dev.off()

@

@@

<<simpleMode,eval=F, echo = F, results = hide>>=

## plotting a hypothesed land cover change model
data1 <- data.frame(year=c(1,2,2,3),tc=c(2,2,1,1))
data2 <- data.frame(year=c(1,2,2,3,4),tc=c(2,2,1,1.5,2))

pdf("tc_simple.pdf", width = 9, height = 4)
par(mar=c(5,6,1,1), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",
     axes=F, frame.plot=F)
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
lines(c(1,4),c(2,2),lty=2)
lines(c(1,4),c(1,1),lty=2)
lines(c(2,2),c(0,2.5),lty=2)
text(2.8,2.1,"(a)", font=2)
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",axes=F, frame.plot=F)
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
lines(c(1,4),c(2,2),lty=2)
lines(c(1,4),c(1,1),lty=2)
lines(c(2,2),c(0,2.5),lty=2)
text(3.5,2.1,"(b)", font=2)
dev.off()

@

<<fullRegression,eval=F , echo = F, results = hide>>=

## predicting model: land cover is category variable
rain.trend <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      var.data <- na.omit(var.data)    
      rain_fit <- gam(rain ~ s(month, k=3)
                  + s(SOI, k=3)
                  + cover  
                  + trd
                  , data=var.data,family=Gamma(link=log))
    } else
     if (grepl("nsw",substitute(y))) {  
      var.data <- na.omit(var.data)    
      rain_fit <- gam(rain ~ s(month, k=3)
                  + s(SOI, k=3)
                  + s(IOD, k=3)
                  + cover  
                  + trd
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}

## note: link=log generally has higher r.sq value than link=reverse

## generate variables for the model
trd <- seq(1:360)
month <-rep(1:12,times=30)
mth <- cycle(ts(month,start=c(1979,1),frequency=12))

LC <- c(rep("tree",288),rep("clear",72))
LC_qld4 <- c(rep("tree",312),rep("clear",48)) ## after- : from Jan 2005

var.data <- data.frame(month, trd, SOI = SOI_30a, IOD = IOD_30a, cover = LC)
var.qld4 <- data.frame(month, trd, SOI = SOI_30a, IOD = IOD_30a, cover = LC_qld4)

  coef.rain30.qld4 <- lapply(2:ncol(qld_30), 
                             function(x) rain.trend(x,qld_30,30,
                                                    var.qld4))
  coef.rain30.nswvic <- lapply(2:ncol(nsw_30), 
                               function(x) rain.trend(x,nsw_30,
                                                      30,var.data))

## extract sign of C
  C30_qld <- unlist(lapply(1:length(coef.rain30.qld4),
    function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                       summary(coef.rain30.qld4[[x]])$p.coeff[2],
                       NA)))
  C30_nswvic <- unlist(lapply(1:length(coef.rain30.nswvic),
    function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                       summary(coef.rain30.nswvic[[x]])$p.coeff[2],
                       NA)))

Cp30_qld4 <- unlist(lapply(1:length(coef.rain30.qld4),
   function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                      summary(coef.rain30.qld4[[x]])$p.pv[2],NA)))
Cp30_nswvic <- unlist(lapply(1:length(coef.rain30.nswvic),
    function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                       summary(coef.rain30.nswvic[[x]])$p.pv[2],
                       NA)))

## p value of seasons and climate indices
p_qld <- unlist(lapply(1:length(coef.rain30.qld4),
      function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                         summary(coef.rain30.qld4[[x]])$s.pv[1],
                         NA)))
p_nsw <- unlist(lapply(1:length(coef.rain30.nswvic),
      function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                         summary(coef.rain30.nswvic[[x]])$s.pv[1],
                         NA)))

## extract r.sq of the model
rsq_qld <- unlist(lapply(1:length(coef.rain30.qld4),
     function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                        summary(coef.rain30.qld4[[x]])$r.sq,NA)))
rsq_nsw <- unlist(lapply(1:length(coef.rain30.nswvic),
     function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                        summary(coef.rain30.nswvic[[x]])$r.sq,NA)))

## frame of the two regions (100 years data)
qld_tif <- raster("Data\\BoM\\rainfall\\AWAP\\from1900_FVG\\r19000101_qld.tif")
nsw_tif <- raster("Data\\BoM\\rainfall\\AWAP\\from1900_FVG\\r19000101_nsw.tif")

  dist_Cp30_qld4 <- raster(qld_tif)
  dist_Cp30_qld4[] <- Cp30_qld4
  dist_Cp30_nsw <- raster(nsw_tif)
  dist_Cp30_nsw[] <- Cp30_nswvic
# ## range of predicted values
  predictmax_qld30 <- unlist(lapply(1:length(coef.rain30.qld4),
        function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",
                          max(coef.rain30.qld4[[x]]$fitted.values),
                           NA)))
  predictmin_qld30 <- unlist(lapply(1:length(coef.rain30.qld4),
        function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",                 min(coef.rain30.qld4[[x]]$fitted.values),NA)))
# 
mean_qld <- lapply(1:length(coef.rain30.qld4), 
          function(x) if(summary(coef.rain30.qld4)[x,2]=="gam") {fitted(coef.rain30.qld4[[x]])
            } else{NA})

## calculate mean residual
  cc <- 1
  qld.sum <- mean_qld[[1]]
  for (i in 2:length(mean_qld)) {
    if(!is.na(mean_qld[[i]])) {
      qld.sum <- qld.sum + mean_qld[[i]]
      cc <- cc+1
    }
  }
  qld.mean <- qld.sum/cc
  qld.mean <- ts(qld.mean,start=c(1979,1),frequency=12)

predictmax_nsw30 <- unlist(lapply(1:length(coef.rain30.nswvic),
        function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                      max(coef.rain30.nswvic[[x]]$fitted.values),
                      NA)))
# 70 - 254.7
  predictmin_nsw30 <- unlist(lapply(1:length(coef.rain30.nswvic),
        function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",
                      min(coef.rain30.nswvic[[x]]$fitted.values),
                      NA)))
# 18.4 - 62.2

mean_nswvic <- lapply(1:length(coef.rain30.nswvic), 
          function(x) if(summary(coef.rain30.nswvic)[x,2]=="gam") {
              fitted(coef.rain30.nswvic[[x]])} else{NA})

## calculate mean residual
  cc <- 1
  nswvic.sum <- mean_nswvic[[1]]
  for (i in 2:length(mean_nswvic)) {
    if(!is.na(mean_nswvic[[i]])) {
      nswvic.sum <- nswvic.sum + mean_nswvic[[i]]
      cc <- cc+1
    }
  }
  nswvic.mean <- nswvic.sum/cc
  nswvic.mean <- ts(nswvic.mean,start=c(1979,1),frequency=12)

## plots
  pdf("Cp_30yrs.pdf", width=10, height=4.5)
  par(mar=c(3,3,1,6),xpd=T, mfrow=c(1,2))
  plot(e111,xlab="",ylab="")
  image(dist_Cp30_qld4,col=c("red","orange","yellow",
                            "light blue"),
        breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  plot(e111,add=T)
text(146, -25.4, "(a)", font=2)
image.plot(zlim=c(0,4),breaks=c(0,1,2,3,4),lab.breaks=c(0,0.05,0.1,0.2,1),legend.only=T,horizontal=F,col=c("red","orange","yellow","light blue"),legend.shrink=0.7,legend.width=0.9,legend.args=list(text="p value",line=0.5,cex=1.2,font=1),axis.args=list(cex.axis=0.8))
legend(148.7,-27.9,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=1)
  plot(e222,xlab="",ylab="") #,border="white")
  image(dist_Cp30_nsw,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
  plot(e222,add=T)
  text(147.1, -34.96, "(b)", font=2)
  dev.off()
# 
# 
## plot gam.check
  pdf("qld_check.pdf")
  gam.check(coef.rain30.qld4[[1]])
  dev.off()

  pdf("nswvic_check.pdf")
  gam.check(coef.rain30.nswvic[[which(values(dist_Cp30_nsw)<0.05)[6]]])
  dev.off()

@

@@

%chunk 5: note code has changed
<<chunk5,eval=F , echo = F, results = hide>>=

## without land cover variable for step rend test
rain.trend3 <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      rain_fit <- gam(rain ~ trd  # 
                  + s(month,k=3)
                  + s(SOI, k=3)
                  , data=var.data,family=Gamma(link=log))
    } else
    if (grepl("nsw",substitute(y))) {
      rain_fit <- gam(rain ~ trd
                  + s(month,k=3)
                  + s(SOI, k=3)
                  + s(IOD, k=3) 
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}
# 
# 
## run model for 30 years
  coef.rain.qld <- lapply(2:ncol(qld_30), 
                function(x) rain.trend3(x,qld_30,30,var.data))
  coef.rain.nswvic <- lapply(2:ncol(nsw_30), 
                function(x) rain.trend3(x,nsw_30,30,var.data))

## explaining power of the partial model
## QLD
  rsq_qld <- unlist(lapply(1:length(coef.rain.qld), 
          function(x) ifelse(summary(coef.rain.qld)[x,2]=="gam",
                             summary(coef.rain.qld[[x]])$r.sq,
                             NA)))
# 
## extract residual
  residual_qld <- lapply(1:length(coef.rain.qld), 
            function(x) if(summary(coef.rain.qld)[x,2]=="gam") {
              coef.rain.qld[[x]]$y-fitted(coef.rain.qld[[x]])
              } else {NA})

# 
# ## NSW/VIC
# 
  rsq_nswvic <- unlist(lapply(1:length(coef.rain.nswvic), 
        function(x) ifelse(summary(coef.rain.nswvic)[x,2]=="gam",
                           summary(coef.rain.nswvic[[x]])$r.sq,
                           NA)))
#

# extract residual
residual_nswvic <- lapply(1:length(coef.rain.nswvic), 
        function(x) if(summary(coef.rain.nswvic)[x,2]=="gam") {
          coef.rain.nswvic[[x]]$y-fitted(coef.rain.nswvic[[x]])
          } else {NA})
saveRDS(residual_nswvic, file="residual_nswvic.RDS")
saveRDS(residual_qld, file="residual_qld.RDS")

## explaining power of the partial model
rain.trend4 <- function(x,y,m,var.data) {

  var.data$rain <- as.vector(unlist(y[x]))+0.0001

  if (!is.na(mean(var.data$rain))) {
    if (grepl("qld",substitute(y))) {
      rain_fit <- gam(rain ~ s(SOI, k=3)
                  , data=var.data,family=Gamma(link=log))
    } else
    if (grepl("nsw",substitute(y))) {
      rain_fit <- gam(rain ~ s(SOI, k=3)
                  + s(IOD, k=3) 
                  , data=var.data,family=Gamma(link=log))
    }

  return(rain_fit)
  } else {NA}

}

## explaining power of index
  coef.soi.qld <- lapply(2:ncol(qld_30), 
                function(x) rain.trend4(x,qld_30,30,var.data))
  coef.soi.nswvic <- lapply(2:ncol(nsw_30), 
                  function(x) rain.trend4(x,nsw_30,30,var.data))
  rsq_soi_qld <- unlist(lapply(1:length(coef.soi.qld), 
          function(x) ifelse(summary(coef.soi.qld)[x,2]=="gam",
                             summary(coef.soi.qld[[x]])$r.sq,NA)))
  rsq_soi_nswvic <- unlist(lapply(1:length(coef.soi.nswvic),
          function(x) ifelse(summary(coef.soi.nswvic)[x,2]=="gam",
                             summary(coef.soi.nswvic[[x]])$r.sq,
                             NA)))

pdf("soi.pdf", width = 10, height = 4)
par(mfrow = c(1,2))
 hist(rsq_soi_qld*100,main="",xlab=expression(paste("% of variability explained (",R^2,")")),cex.lab=1.2)
 text(4.2,400, "(a)", font=2)
 hist(rsq_soi_nswvic*100,main="",xlab=expression(paste("% of variability explained (",R^2,")")),cex.lab=1.2)
text(3.2,300, "(b)", font=2)
dev.off()

@

@@

%chunk 6
<<chunk6,eval=F , echo = F, results = hide>>=

## perform step test on the data
## function to calculate the step trend statistic (Hirsch, 1985)
stepTrend2 <- function(x,y,i){  
    n1 <- 2002-1997
    if (i==1){
      z <- 56  # no. of column
    } else {
    z <- 37 } # close if

  if (x%%z!=0){ #& x%%z!=(z-1)) {
    y.1 <- y[[x]]
    y.2 <- y[[x+1]]
    y.3 <- y[[x+z]]
    y.4 <- y[[x+z+1]]

  if(!is.na(y.1[1])&!is.na(y.2[1])&!is.na(y.3[1])&!is.na(y.4[1])) {
    if (i == 1) {
      n2 <- 2008-2004    
      v.1 <- ts(y.1[c(229:288,313:360)],start=c(1998,1),
                frequency=12)
      v.2 <- ts(y.2[c(229:288,313:360)],start=c(1998,1),
                frequency=12)
      v.3 <- ts(y.3[c(229:288,313:360)],start=c(1998,1),
                frequency=12)
      v.4 <- ts(y.4[c(229:288,313:360)],start=c(1998,1),
                frequency=12)
    } else {
      n2 <- 2008-2002
      v.1 <- ts(y.1[c(229:360)],start=c(1998,1),frequency=12)
      v.2 <- ts(y.2[c(229:360)],start=c(1998,1),frequency=12)
      v.3 <- ts(y.3[c(229:360)],start=c(1998,1),frequency=12)
      v.4 <- ts(y.4[c(229:360)],start=c(1998,1),frequency=12)
    } # close if (i ==1)
  #browser()
## rank by month
   for (j in 1:12) { 
    assign(paste("r1",j,sep=""),rank(subset(v.1,cycle(v.1)==j)))
    assign(paste("r2",j,sep=""),rank(subset(v.2,cycle(v.2)==j)))
    assign(paste("r3",j,sep=""),rank(subset(v.3,cycle(v.3)==j)))
    assign(paste("r4",j,sep=""),rank(subset(v.4,cycle(v.4)==j))) 
   } # close rank by month
## mann-whitney rank sum statistic
    rlist.1 <- list(r11,r12,r13,r14,r15,r16,r17,r18,r19,r110,
                    r111,r112) ## full year
    rlist.2 <- list(r21,r22,r23,r24,r25,r26,r27,r28,r29,r210,
                    r211,r212)
    rlist.3 <- list(r31,r32,r33,r34,r35,r36,r37,r38,r39,r310,
                    r311,r312)
    rlist.4 <- list(r41,r42,r43,r44,r45,r46,r47,r48,r49,r410,
                    r411,r412)
    rlist <- list(unlist(rlist.1),unlist(rlist.2),unlist(rlist.3),unlist(rlist.4)) 
    w.s1 <- 0
    w.s2 <- 0
    w.s3 <- 0
    w.s4 <- 0
  for (k in 1:length(rlist.1)) {
    w.s1 <- sum(rlist.1[[k]][1:n1])+w.s1
    w.s2 <- sum(rlist.2[[k]][1:n1])+w.s2
    w.s3 <- sum(rlist.3[[k]][1:n1])+w.s3
    w.s4 <- sum(rlist.4[[k]][1:n1])+w.s4
    }
    w.s <- sum(w.s1,w.s2,w.s3,w.s4) #,w.s5,w.s6,w.s7,w.s8,w.s9)

>>>>

\usepackage{Sweave}
\begin{document}
\input{chap3_paper_new-concordance}

@@

@@

@@

@@

@@

%chunk 5: note code has changed

@@

%chunk 6

% chunk 8: bootstrapping (spatial structure is kept, used)

% chunk 9: bootstrapping (spatial structure is kept, remove short term trend, randomly select after change period)
% apply the LocSig function in SpatialVx for block bootstrapping
